---
title: "scatter"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(here)
library(tidyverse)
library(patchwork)
```


```{r}
indicators <- read_csv("condata.csv")[,1:5] 
  #mutate(indicator = tolower(indicator),#this made all names lower case 
#read_csv gewts rid of trailing and white space 
        # group = ifelse(group == "water quality", "Water Quality", as.character(group))) #there was two water quality groups so I change them to be just one now.

#hashed these out because i fixed it in excel when doing quality check 

ggplot(indicators, aes(x=group, y=mentions, fill= indicator)) +
  geom_col() + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap (vars(method), scales = "free_y", nrow = 2) #automatically does same scale to change this adding free scales argment. 
#nrow is 1 = next to each other, 2 = stacked 


```
```{r}
# Just wondering - why are the values here between 0 and 1? Does the y axis represent the proportion of mentions of each type or??
ggplot(indicators, aes(x=group, y=mentions, fill= indicator)) +
  geom_col(position = "fill") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap (vars(method), scales = "free_y", nrow = 2) 
```



```{r}
monitor<-indicators %>% 
  filter(method == "Monitoring")

mon_plot<-ggplot(monitor, aes(x=indicator, y=mentions, fill= method)) +
  geom_bar(position="stack", stat="identity")+ #changed to stack bar so that you can see the # of mentions
  theme_bw()+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap (vars(type), nrow = 3) +
  coord_polar(start = 0)


mon_plot
```


```{r}
msp<-indicators %>% 
  filter(method == "MSP")

msp_plot<-ggplot(msp, aes(x=group, y=mentions, fill= indicator)) +
  geom_bar(position="stack", stat="identity")+ 
    theme_bw()+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap (vars(type), nrow = 3) 

msp_plot
```
```{r}
mon_plot + msp_plot
```


#for the scatterplot re-shape:
```{r}
library(tidyr)

scatterindi<- indicators %>% 
  pivot_wider(names_from = method, values_from = mentions) %>% 
  replace_na(list(MSP = 0, Monitoring = 0)) 
head(scatterindi)

```

```{r}
ggplot(indicators, aes(x=indicator, y=mentions, fill= method)) +
 geom_bar(position="stack", stat="identity")+ 
    theme_bw()+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  coord_polar(start = 0)

```


```{r}
  ggplot(scatterindi, aes(x= MSP, y= Monitoring, colour = group)) + 
  geom_point(aes(shape = type))

#i had to convert the NA to '0' so its dense there, maybe i should do a thing about this? 
#i want to do something [maybe add a circle around it?] if the indicator is one that is both MSP and monitoring, unsure of how~!

#position= figure this out to spepearte dots when overlapping
  #count, point, jitter to display it all 

  
```

```{r}
monitor<-scatterindi %>% 
  filter(method == "Monitoring")

mon_plot<-ggplot(monitor, aes(x=group, y=mentions, fill= indicator)) +
  geom_bar(position="stack", stat="identity")+ #changed to stack bar so that you can see the # of mentions
  theme_bw()+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap (vars(type), nrow = 3) 
```




# Trying a heat/tile map

```{r}
tile_plot<-ggplot(indicators, aes(method, indicator, fill = mentions)) +
  geom_tile()+
  scale_fill_gradient(low = "white", high = "blue") +
  facet_wrap(~type)

tile_plot

# You could work on ordering the "indicators" on the y-axis into groups based on the "group" column and labelling that on the y-axis to make it clear what braod categories each fall into
```

Try plotting them separately bc obviously there isn't overlap between the biological, physical and socio-economic indicators
```{r}
biol<-indicators %>% 
  filter(type == "Biological")

biol_plot<-ggplot(biol, aes(method, indicator, fill = mentions)) +
  geom_tile()+
  scale_fill_gradient(low = "white", high = "blue") +
  facet_wrap(~type)

biol_plot
```

```{r}
phys<-indicators %>% 
  filter(type == "Physical")

phys_plot<-ggplot(phys, aes(method, indicator, fill = mentions)) +
  geom_tile()+
  scale_fill_gradient(low = "white", high = "blue") +
  facet_wrap(~type)

phys_plot
```

```{r}
social<-indicators %>% 
  filter(type == "Socio-economic")

soc_plot<-ggplot(social, aes(method, indicator, fill = mentions)) +
  geom_tile()+
  scale_fill_gradient(low = "white", high = "blue") +
  facet_wrap(~type)

soc_plot
```

```{r}
# this functionality for arranging plots if from the patchwork package
# see here for help/more info https://gotellilab.github.io/GotelliLabMeetingHacks/NickGotelli/ggplotPatchwork.html

biol_plot + phys_plot + soc_plot + plot_layout(ncol = 1)

# I think I like this if you can then add the groupings of the "group" types along the y-axis as well (you would need to reorder the y-axis names based on the "group" then could add in the group in powerpoint or illustrator afterwards)
```


# Trying a spider chart

```{r}
library(fmsb)

# Create data: note in High school for several students
set.seed(99)
data <- as.data.frame(matrix( sample( 0:20 , 15 , replace=F) , ncol=5))
colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding" )
rownames(data) <- paste("mister" , letters[1:3] , sep="-")
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each variable to show on the plot!
data <- rbind(rep(20,5) , rep(0,5) , data)

spider_data<-indicators %>% 
  group_by(type, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "type", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data<-rbind(rep(254,3), rep(0, 3),spider_data)

radarchart(spider_data)

# Tutorial to make it prettier here: https://www.r-graph-gallery.com/143-spider-chart-with-saveral-individuals.html

spider_data2<-indicators %>% 
  group_by(group, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "group", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data2[is.na(spider_data2)]<-0

spider_data2<-rbind(rep(145,13), rep(0, 13),spider_data2)

radarchart(spider_data2)

# Could do two separate ones - one for MSP an done for monitoring to show overlap of biol, phys and socio indicators if there is any
```
```{r}

bio<-indicators %>% 
  filter(type == "Biological")

spider_data3<-bio %>% 
  group_by(group, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "group", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data3[is.na(spider_data3)]<-0

spider_data3<-rbind(rep(145,13), rep(0, 13),spider_data3)

radarchart(spider_data3)
```
```{r}

phys<-indicators %>% 
  filter(type == "Physical")

spider_data4<-phys %>% 
  group_by(group, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "group", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data4[is.na(spider_data4)]<-0

spider_data4<-rbind(rep(145,13), rep(0, 13),spider_data4)

radarchart(spider_data4)
```
this one dosent work because too low indicators 
#

soc<-indicators %>% 
  filter(type == "Socio economic")

spider_data5<-soc %>% 
  group_by(indicator, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "indicator", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data5[is.na(spider_data5)]<-0

spider_data5<-rbind(rep(145,13), rep(0, 13),spider_data5)

radarchart(spider_data5)
```{r}
indicators
soc<-indicators %>% 
  filter(type == "Socio-economic")

spider_data5<-soc %>% 
  group_by(group, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "group", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data5[is.na(spider_data5)]<-0

spider_data5<-rbind(rep(10,88), rep(0, 88),spider_data5)

radarchart(spider_data5)
```


```{r}

phys<-indicators %>% 
  filter(type == "Physical")

spider_data5<-phys %>% 
  group_by(indicator, method) %>% 
  summarise(total_m = sum(mentions, na.rm = T)) %>% 
  pivot_wider(., names_from = "indicator", values_from = "total_m") %>% 
  column_to_rownames(., var = "method")

spider_data5[is.na(spider_data5)]<-0

spider_data5<-rbind(rep(14,13), rep(0, 13),spider_data5)

#the rbind is (rep())

radarchart(spider_data5)
```